Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean (*Phaseolus vulgaris*) in the United States

### 2017-11-02 doc/manuscript/manuscript.Rmd
update visualization a bit
fix P-value format and add citation for protocols
These address #15
I also added Reviewer 2's comment on the introduction

*Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen with a worldwide distribution [@bolton2006sclerotinia].
This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) and has the ability to survive for more than five years in soil using melanized survival structures called sclerotia [@bolton2006sclerotinia; @sexton2006population]. 
It causes disease on more than 400 plant species belonging to 75 families [@boland1994index] including crops of major economic importance such as sunflower (*Helianthus spp.*), soybean (*Glycine max* L.), canola (*Brassica napa* L., *Brassica campestris* L.), and dry bean (*Phaseolus vulgaris* L.) [@bolton2006sclerotinia]. 
On dry bean, *S. sclerotiorum* is the causal agent of white mold, a devastating disease that can be yield-limiting in temperate climates [@steadman1983white].

Currently, there are no commercially available resistant cultivars of dry bean [@otto-hanson2011variation].
Organized breeding efforts have used a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Australia, France, and Mexico [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use]. 
These white mold screening nurseries use no chemical or cultural treatments against *S. sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold. 
It has previously been shown that aggressiveness (the severity of disease symptoms on the host) is significantly different across white mold screening nursery sites in separate geographic regions [@otto-hanson2011variation].
The genetic structure and mode of reproduction in these populations, however, is currently unknown.

@@

## Data processing and Analysis{-}

@@

loc_table %>%
  mutate(allele = round(allele, 2)) %>%
  mutate(Repeats = case_when(
    Locus == "5-2"   ~ "(GT)",
    Locus == "6-2"   ~ "(TTTTTC)(TTTTTG)(TTTTTC)", # "(TTTTTC)~2~(TTTTTG)~2~(TTTTTC)",
    Locus == "7-2"   ~ "(GA)",
    Locus == "8-3"   ~ "(CA)",
    Locus == "9-2"   ~ "(CA)(CT)", # "(CA)~9~(CT)~9~",
    Locus == "12-2"  ~ "(CA)",
    Locus == "17-3"  ~ "(TTA)",
    Locus == "20-3"  ~ "(GT)GG(GT)", # "(GT)~7~GG(GT)~5~",
    Locus == "55-4"  ~ "(TACA)",
    Locus == "110-4" ~ "(TATG)",
    Locus == "114-4" ~ "(TAGA)"
  )) %>%
  rename(`*h*` = Hexp) %>%
  rename(`No. alleles` = allele) %>%
  rename(`Repeat Motif` = Repeats) %>%
  huxtable::as_huxtable(add_colnames = TRUE) %>%
  huxtable::set_number_format(huxtable::everywhere, 3:4, 0) %>%
  huxtable::set_number_format(huxtable::everywhere, 5:6, 2) %>%
  huxtable::set_align(huxtable::everywhere, 2:6, "right") %>%
  huxtable::set_col_width(c(0.06, 0.07, 0.25, 0.1, 0.05, 0.075)) %>%
  huxtable::print_md(max_width = 90)
```

@@

Agressiveness of the isolates ranged from `r paste(range(other(dat11)$meta$Severity), collapse = " to ")` with a mean of `r signif(mean(other(dat11)$meta$Severity), 3)` and median of `r signif(median(other(dat11)$meta$Severity), 3)`.
The group mean averages were `r sprintf("%0.2f, %0.2f, and %0.2f", Raggmean, Maggmean, Haggmean)` for Region, MCG, and MLH, respectively.
Our assessment of aggressiveness in association with Region showed a significant effect (*P* < 1.00e^-4^), with means that ranged from 5.8 (MN) to 4.0 (CA) (Fig. \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)). 
MCGs also showed a significant effect (*P* < 1.00e^-4^), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49'; Table \@ref(tab:MCG-aggressiveness)).
We additionally found a significant effect for MLHs (*P* = 7.44e^-4^), with means that ranged from 6.0 ('MLH 78') to 4.3 ('MLH 140') (Table \@ref(tab:MLG-aggressiveness)).

>>>>

<!-- 
Reviewer 2:
I would expand introduction a bit to describe the symptoms in greater 
details, maybe even provide some information regarding economic damage (if 
available) to nursery production. I was wondering what cultivars previous 
studies focused on – the authors provided 3 cultivars at the end of 
introduction. Maybe adding another paragraph on differences among 
cultivars – they chose 3 with different levels of resistance but how was 
that selected (assuming some preliminary data)? For someone not familiar 
with the host, I would like to have more details on both host and pathogen. 
-->
*Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen with a worldwide distribution [@bolton2006sclerotinia].
This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) and has the ability to survive for more than five years in soil using melanized survival structures called sclerotia [@bolton2006sclerotinia; @sexton2006population]. 
It causes disease on more than 400 plant species belonging to 75 families [@boland1994index] including crops of major economic importance such as sunflower (*Helianthus spp.*), soybean (*Glycine max* L.), canola (*Brassica napa* L., *Brassica campestris* L.), and dry bean (*Phaseolus vulgaris* L.) [@bolton2006sclerotinia]. 
On dry bean, *S. sclerotiorum* is the causal agent of white mold, a devastating disease that can be yield-limiting in temperate climates [@steadman1983white].

Currently, there are no commercially available resistant cultivars of dry bean [@otto-hanson2011variation].
Organized breeding efforts have used a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Australia, France, and Mexico [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use]. 
These white mold screening nurseries use no chemical or cultural treatments against *S. sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold [@steadman2003identification; @otto-hanson2011variation].
It has previously been shown that aggressiveness (the severity of disease symptoms on the host) is significantly different across white mold screening nursery sites in separate geographic regions [@otto-hanson2011variation].
The genetic structure and mode of reproduction in these populations, however, is currently unknown.

@@

## Data processing and Analysis {-}

@@

loc_table %>%
  mutate(allele = round(allele, 2)) %>%
  mutate(Repeats = case_when(
    Locus == "5-2"   ~ "(GT)",
    Locus == "6-2"   ~ "(TTTTTC)(TTTTTG)(TTTTTC)", # "(TTTTTC)~2~(TTTTTG)~2~(TTTTTC)",
    Locus == "7-2"   ~ "(GA)",
    Locus == "8-3"   ~ "(CA)",
    Locus == "9-2"   ~ "(CA)(CT)", # "(CA)~9~(CT)~9~",
    Locus == "12-2"  ~ "(CA)",
    Locus == "17-3"  ~ "(TTA)",
    Locus == "20-3"  ~ "(GT)GG(GT)", # "(GT)~7~GG(GT)~5~",
    Locus == "55-4"  ~ "(TACA)",
    Locus == "110-4" ~ "(TATG)",
    Locus == "114-4" ~ "(TAGA)"
  )) %>%
  rename(`*h*` = Hexp) %>%
  rename(`No. alleles` = allele) %>%
  rename(`Repeat Motif` = Repeats) %>%
  huxtable::as_huxtable(add_colnames = TRUE) %>%
  huxtable::set_number_format(huxtable::everywhere, 3:4, 0) %>%
  huxtable::set_number_format(huxtable::everywhere, 5:6, 2) %>%
  huxtable::set_align(huxtable::everywhere, 2:6, "right") %>%
  huxtable::set_col_width(c(0.06, 0.07, 0.25, 0.1, 0.05, 0.075)) %>%
  huxtable::print_md(max_width = 100)
```

@@

Agressiveness of the isolates ranged from `r paste(range(other(dat11)$meta$Severity), collapse = " to ")` with a mean of `r signif(mean(other(dat11)$meta$Severity), 3)` and median of `r signif(median(other(dat11)$meta$Severity), 3)`.
The group mean averages were `r sprintf("%0.2f, %0.2f, and %0.2f", Raggmean, Maggmean, Haggmean)` for Region, MCG, and MLH, respectively.
Our assessment of aggressiveness in association with Region showed a significant effect (*P* < 1.00e^-4^), with means that ranged from 5.8 (MN) to 4.0 (CA) (Fig. \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)). 
MCGs also showed a significant effect (*P* < 0.001), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49'; Table \@ref(tab:MCG-aggressiveness)).
We additionally found a significant effect for MLHs (*P* < 0.001), with means that ranged from 6.0 ('MLH 78') to 4.3 ('MLH 140') (Table \@ref(tab:MLG-aggressiveness)).
